---
title: "2019 Karluk Steelhead preliminary estimate"
author: "Adam Reimer"
date: "December 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(magrittr)
library(ggplot2)
library(aslpack)
load("..\\data\\dat19.rda")
```

# Abundance estimate  
  
## Size Selective Sampling Tests  
With respect to size there is no evidence of selective sampling during either event. None of these plots look great to me but the test statistic is insignificant for all three tests. These tests all point toward used of a pooled Chapman estimator.
```{r echo = FALSE}
test_ks <- function(lg1, lg2, title){
  name1 <- deparse(substitute(lg1))
  name2 <- deparse(substitute(lg2))
  plot(ecdf(lg1),
       main = title,
       ylab = "Proportion",
       xlab = "length")
  plot(ecdf(lg2), add = TRUE, col = "red")
  legend(quantile(lg1, .65, na.rm = TRUE), .40, legend = c(name1, name2), col = c("black", "red"), pch = 16)
  test <- ks.test(as.numeric(lg1),
                  as.numeric(lg2))
  text(quantile(lg1, .65, na.rm = TRUE), .1,
       labels = paste0(test$method, "\n p.val = ", format.pval(test$p.value, digits = 2, eps = 0.001, nsmall = 3)),
       adj = 0)
}
```

```{r}
#no size selectivity
R_lg <- dat19$R$length[!is.na(dat19$R$length)]
C_lg <- dat19$C$length[!is.na(dat19$C$length)]
M_lg <- dat19$M$length[!is.na(dat19$M$length)]
test_ks(C_lg, R_lg, "C vs. R")
test_ks(M_lg, R_lg, "M vs. R")
test_ks(M_lg, C_lg, "M vs. C")
```
  
## Sex Selective Sampling Tests
With respect to sex selective sampling we fall into case V (inconsistent test results). The issue is that both sex selectivity tests point to non-selective sampling but the sex composition of the two events differs. Two interpretations are possible. First, we could assume the sex comp at the weir is the better estimate and the event specific tests likely have low power due to a small sample sizes. Under this interpretation the mark-recapture can still be pooled but we would have to be careful about using the age/sex composition stuff from the marking event because we believe that sampling may have been selective. The other interpretation is that both events had non-selective sampling but that survival differed between events.
```{r}
#sex selectivity
#first event unbiased
R_sex <- as.character(dat19$R$sex[!is.na(dat19$R$sex)])
C_sex <- as.character(dat19$C$sex[!is.na(dat19$C$sex)])
M_sex <- as.character(dat19$M$sex[!is.na(dat19$M$sex)])
CR <- data.frame(strata= c(rep("C", length(C_sex)), rep("R", length(R_sex))),
                 sex = c(C_sex, R_sex))
aslpack::tab_lr(CR, "sex")

#second event unbiased
MR <- data.frame(strata= c(rep("M", length(M_sex)), rep("R", length(R_sex ))),
                 sex = c(M_sex, R_sex))
aslpack::tab_lr(MR, "sex")

MC <- data.frame(strata= c(rep("M", length(M_sex)), rep("C", length(C_sex ))),
                 sex = c(M_sex, C_sex))
aslpack::tab_lr(MC, "sex", )
```
  
## Consistency tests
Consistency tests involve whether temporal stratification is required. All tests were all passed, pointing towards the use of a Peterson estimator.
```{r}
#Consistancy tests passed
mix_tab <- cbind(table(dat19$R$mark_strata, dat19$R$strata), table(dat19$M$strata) - table(dat19$R$mark_strata))
mix_tab
DescTools::GTest(mix_tab)

eqprop_tab <- rbind(table(dat19$R$strata), table(dat19$C$strata) - table(dat19$R$strata))
eqprop_tab
DescTools::GTest(eqprop_tab)

completemix_tab <- rbind(table(dat19$R$mark_strata), table(dat19$M$strata) - table(dat19$R$mark_strata))
completemix_tab
chisq.test(completemix_tab)
```

The strength of this project is the idea that the weir is a kelt census satisfying most of the typical mark-recapture assumptions. You expressed doubts about the weir as a kelt census and that thought is defiantly supported in some years. That said, I don't see any obvious truncation this year.  
```{r}
#weir captures whole run
ggplot(dat19$weir, aes(x = as.Date(date), weight = daily)) +
  geom_histogram(binwidth = 1) +
  scale_x_date(date_labels = "%m/%d", date_breaks = "weeks") +
  labs(x = "Date")
```

Here are the relevant sample sizes, the estimated abundance and the SE of the estimated abundance.
```{r, echo = TRUE}
M <- dim(dat19$M)[1]; M
C <- max(dat19$weir$cum)[1]; C
R <- sum(dat19$R$Ytag); R
N <- (M + 1) * (C + 1) / (R + 1) - 1; N
N_se <- sqrt((M + 1) * (C + 1) * (M - R) * (C - R) / (R + 1) ^2 / (R + 2)); N_se
```
  
## ASL at the weir
Timing of the ASL samples was not representative of emigration in general...
```{r}
#But... ASL samples are not representative of emmigration in general
plot_w <- dat19$weir[, c("date", "daily")]
plot_w$source = "Weir count"

plot_age <-
  aggregate(lg_type ~ date, dat19$C, FUN = length) %>%
  setNames(c("date", "daily"))
plot_age$source <- "age samples"

dplyr::left_join(plot_w, plot_age, by = "date") %>%
  dplyr::select(date,
                emigration = daily.x,
                samples = daily.y) %>%
  dplyr::mutate(samples = ifelse(is.na(samples), 0, samples)) %>%
  aslpack::plot_ks("emigration")
```
  
But sex is time invariant. The rows represent temporal strata at the weir (I only chose two strata because the main differences between sampling and emigration fell on either side of the 50% cumulative passage mark).
```{r}
#But... Sex and age comp are time invariant
#Redefine strata since major difference are early and late
C19 <- dat19$C #new dataframe since aslpack::tab_lr looks for a varible named "strata"
C19$strata <- cut(dat19$C$date,
                       quantile(rep(dat19$weir$date, times = dat19$weir$daily), c(0, .5, 1)),
                       labels = FALSE)
aslpack::tab_lr(C19[!is.na(C19$sex), ], "sex")
```
  
Saltwater age is also time invariant.
```{r}
aslpack::tab_lr(C19[!is.na(C19$age_s), ], "age_s")
```

Here is the ASL table for kelts at the weir. I followed the convention from earlier reports where the estimates were separated into saltwater age and initial/repeat spawners. In the early reports they had repeat and multi-repeat categories but we only had 3 fish that were multi-repeat. This table is a little messy but will look nicer with some excel formatting.
```{r}
#Convert aslpack::tab_asl to accept non-euro ages
test <-
  function(dat,
           totalname = "Total",
           output = "asl",
           display_cols = NULL,
           gather_cols = c("n.z", "p.z", "ci_p.z", "t.z", "lg.z", "range_lg.z"),
           age_labels = NULL,
           overall_se = NULL){
    stopifnot(!(display_cols %in% c(NULL, "stats")))

    if(!is.null(overall_se)) dat$sd_t.z[dat$age == "All" & dat$sex == "Both"] = overall_se
    if(is.null(age_labels)) age_labels <- c(unique(dat$age)[-length(unique(dat$age))], "All ages")

    temp <-
      dat %>%
      dplyr::mutate_at(c("p.z", "sd_p.z", "lci_p.z", "uci_p.z"), list(~ trimws(format(round(., 3), nsmall = 3)))) %>%
      dplyr::mutate_at(c("lg.z", "se_lg.z"), list(~ trimws(format(round(., 1), big.mark = ",", nsmall = 1)))) %>%
      dplyr::mutate_at(c("n.z", "min_lg.z", "max_lg.z"), as.integer) %>%
      dplyr::mutate(p.z = paste0(p.z, " (", ifelse(sex =="Both" & age == "All", 0, sd_p.z), ")"),
                    ci_p.z = ifelse(sex =="Both" & age == "All", "", paste0(lci_p.z, " - ", uci_p.z)),
                    t.z = paste0(trimws(format(round(t.z, 0), big.mark = ",", nsmall = 0)),
                                 " (",
                                 trimws(format(round(sd_t.z, 0), big.mark = ",", nsmall = 0)),
                                 ")"),
                    lg.z = paste0(lg.z, " (", se_lg.z, ")"),
                    range_lg.z = paste0("(", min_lg.z, "-", max_lg.z, ")")) %>%
      dplyr::select(sex, age, gather_cols) %>%
      tidyr::gather_("measure", "value", gather_cols = gather_cols) %>%
      dplyr::mutate(sex_lab = factor(sex, levels = c("F",       "M",     "Both"),
                                     labels = c("Females", "Males", "Both Sexes"),
                                     ordered = TRUE, exclude = NULL),
                    stat_lab = factor(measure, levels = c("n.z",
                                                          "p.z",
                                                          "ci_p.z",
                                                          "t.z",
                                                          "lg.z",
                                                          "range_lg.z"),
                                      labels = c("Sample size",
                                                 "Proportion (SE)",
                                                 "95% CI(Proportion)",
                                                 paste0(totalname, "(SE)"),
                                                 "Mean Length (SE)",
                                                 "Range(Length)"),
                                      ordered = TRUE),
                    age_lab = factor(age,
                                     levels = unique(dat$age),
                                     labels = age_labels,
                                     ordered = TRUE, exclude = NULL)) %>%
      dplyr::select(sex_lab, stat_lab, age_lab, value)

    switch(output,
           "al" = temp %>%
             dplyr::filter(sex_lab == "Both Sexes") %>%
             dplyr::select(-sex_lab) %>%
             tidyr::spread(if(is.null(display_cols)) age_lab else(stat_lab), value),
           "sl" = temp %>%
             dplyr::filter(age_lab == "All ages") %>%
             dplyr::select(-age_lab) %>%
             tidyr::spread(if(is.null(display_cols)) sex_lab else(stat_lab), value),
           "asl" = temp %>% tidyr::spread(if(is.null(display_cols)) age_lab else(stat_lab), value))
  }


#Sex comp first since unaged fish skew sex comp
sl_tab <-
  dat19$C %>%
  dplyr::select(age = age_s, sex, length) %>%
  dplyr::filter(!is.na(sex)) %>%
  asl(data.frame(total = C, se_total = 0))

weir_sex <- #Emmigration by sex
  sl_tab[sl_tab$age %in% "All", ] %>%
  dplyr::select(sex_strata = sex, total = t.z, se_total = sd_t.z)

age_clean <- #clean data with sex as a stratification varible
  dat19$C %>%
  dplyr::select(age = age_s, sex, length, spawn) %>%
  dplyr::filter(!is.na(age) & !is.na(sex)) %>%
  dplyr::mutate(age2 = ifelse(spawn %in% 1, paste0(age, "-Initial"), paste0(age, "-Repeat")))
age_clean$sex_strata = age_clean$sex

# asl(age_clean, weir_sex, c("sex_strata")) %>%
#   combine_strata() %>%
#   test(totalname = "Kelts", output = "asl", overall_se = 0)

age_clean2 <- age_clean #new dataset which combines age and inital/repeat spawning info in age label
age_clean2$age <- age_clean2$age2
asl(age_clean2, weir_sex, c("sex_strata")) %>%
  combine_strata() %>%
  test(totalname = "Kelts", output = "asl", overall_se = 0) %>%
  knitr::kable()
```
  
## ASL on the spawning grounds 
For the spawning steelhead ASL we are assuming a single representative grab sample. Consider this table a placeholder dependent on your interpretation of the test results. Are you bothered by having fewer age categories in this sample vs the emigrant sample? I think we could do this a few ways depending on how we want to interpret our test results. Recall the mark and recapture samples are not different buy size but are different by sex. I also tested for differences by saltwater age composition and composition of initial/repeat spawners (both were not significant). Do you think the sex composition difference reflects sampling bias from the first event of survival differences?
```{r}
#ASL on spawning grounds
dat19$M %>%
  dplyr::select(age0 = age_s, sex, length, spawn) %>%
  dplyr::filter(!is.na(age0) & !is.na(sex)) %>%
  dplyr::mutate(age = ifelse(spawn %in% 1, paste0(age0, "-Initial"), paste0(age0, "-Repeat"))) %>%
  asl(data.frame(total = N, se_total = N_se)) %>%
  test(totalname = "Spawners", output = "asl") %>%
  knitr::kable()
```
  
## Spawning survival
The first row of numbers is the estimate, the second row of numbers is the SE. I made seperate calculations for initial/repeat spawners because they did that in FDS 97. That said the estimates have widely overlapping CI and we could present one number or by sex.
```{r}
#Spawning survival
#need to expand recaps for sampling fraction
psamp <- sum(!is.na(dat19$R$sex)) / sum(dat19$R$Ytag)
V_psamp <- psamp * (1 - psamp) / (sum(dat19$R$Ytag) - 1)

Rspawn_samp <- table(stringr::str_count(dat19$R$age, "s"))
Rspawn_obs <- Rspawn_samp / psamp
#From taylor series expansion. V(R/S) ~ varR/muS^2 -2muR*covRS/muS^3 +muR^2*varS/muS^4
V_Rspawn_obs <- Rspawn_samp^2 / psamp^4 * V_psamp
Mspawn_obs <- c(sum(dat19$M$spawn %in% "1"), sum(dat19$M$spawn %in% c("2", "3")))
(Rspawn_obs / Mspawn_obs) %>% setNames(c("Initial", "Repeat"))
sqrt(V_Rspawn_obs / Mspawn_obs^2) %>% setNames(c("Initial", "Repeat"))
```
  
## Misc. plots
I replicated a few plots from the prior reports and made a new one of my own. You can decide if any of these are necessary and if so I can clean them up or make some edits as appropriate. This is intended to mimic Figure 6 from FDS 97-6.
```{r, fig.width=10}
#Replicate figure 6 from FDS 97-6
plot_w <- dat19$weir[, c("date", "daily")]
plot_w$source = "Weir count"

plot_R <-
  aggregate(Ytag ~ date, dat19$R, FUN = sum) %>%
  setNames(c("date", "daily"))
plot_R$source <- "recaps"

dplyr::left_join(plot_w, plot_R, by = "date") %>%
  dplyr::mutate(unmarked = daily.x  - ifelse(is.na(daily.y), 0, daily.y)) %>%
  dplyr::select(date, marked = daily.y, unmarked) %>%
  tidyr::gather(group, daily, -date) %>%
  dplyr::mutate(week = gsub("[0-9]+-([0-9]+-[0-9]+)", "\\1", cut(date, "weeks"))) %>%
  dplyr::group_by(week, group) %>%
  dplyr::summarise(weekly = sum(daily, na.rm = TRUE)) %>%
  ggplot(aes(x = week, y = weekly, fill = group)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values = c("marked" = "black", "unmarked" = "gray"))
```

This is intended to mimic Figure 8 from FDS 97-6. The numbers above each bar represent the sample size in each week.
```{r}
#Replicate figure 8 from FDS 97-6
weekly_sex <-
  dat19$C %>%
  dplyr::filter(!is.na(sex)) %>%
  dplyr::mutate(week = gsub("[0-9]+-([0-9]+-[0-9]+)", "\\1", cut(date, "weeks"))) %>%
  dplyr::group_by(week, sex) %>%
  dplyr::summarise(weekly = length(sex))

ggplot(weekly_sex, aes(x = week, y = weekly, fill = sex)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity") +
  scale_fill_manual(values = c("M" = "black", "F" = "gray")) +
  geom_text(data = aggregate(weekly ~ week, weekly_sex, FUN = sum),
            aes(x = week, y = 1, label = weekly),
            inherit.aes = FALSE,
            vjust = -.25)
```

Finally recall we did not see size differences between the events. These plots combine all of the ASL samples from both events. Length distribution by total age.
```{r, fig.height=8, fig.width=10}
#Length composition by total age
# dat19$M %>%
#   dplyr::filter(!is.na(age_f) & !is.na(length)) %>%
#   dplyr::mutate(lg_g = cut(length, breaks = seq(0, 1000, 25))) %>%
#   ggplot(aes(x = lg_g, fill = spawn)) +
#   geom_bar() +
#   facet_grid(age_t ~ .)
#
# dat19$C %>%
#   dplyr::filter(!is.na(age_f) & !is.na(length)) %>%
#   dplyr::mutate(lg_g = cut(length, breaks = seq(0, 1000, 25))) %>%
#   ggplot(aes(x = lg_g, fill = spawn)) +
#   geom_bar() +
#   facet_grid(age_t ~ .)

rbind(dat19$C[, c("length", "age_f", "age_s", "age_t", "spawn")], dat19$M[, c("length", "age_f", "age_s", "age_t", "spawn")]) %>%
  dplyr::filter(!is.na(age_f) & !is.na(length)) %>%
  dplyr::mutate(lg_g = cut(length, breaks = seq(0, 1000, 25))) %>%
  ggplot(aes(x = lg_g, fill = spawn)) +
  geom_bar() +
  facet_grid(age_t ~ .)
```
  
Length distribution by saltwater age.  
```{r, fig.height=8, fig.width=10}
rbind(dat19$C[, c("length", "age_f", "age_s", "age_t", "spawn")], dat19$M[, c("length", "age_f", "age_s", "age_t", "spawn")]) %>%
  dplyr::filter(!is.na(age_s) & !is.na(length)) %>%
  dplyr::mutate(lg_g = cut(length, breaks = seq(0, 1000, 25))) %>%
  ggplot(aes(x = lg_g, fill = spawn)) +
  geom_bar() +
  facet_grid(age_s ~ .)
```

# 2017 abundance
I think we need to revisit the 2017 abundance estimate. I made some pretty strong assumptions that do not look believable in light of the new data. Also, Mark implied the 2017 data may have the same problem as the 2018 data (The same rouge technician). If this is true the best thing would be to drop the 2017 estimate entirely. We should discuss this by phone.
